\clearpage

\section{Identifiability and inference of the individual infectiousness profile's shape}

While the concentration of the individual infectiousness profile can impact how an epidemic unfolds, we lack frameworks for estimating it. Many approaches exist for estimating the generation interval, but there has not been a systematic approach developed to assess how individual infectiousness varies and to specify where on the continuum we sit---i.e., how much of the variation in the generation interval distribution is attributable to variation between individuals, and how much to variation in infection timing within individuals.

We begin by assessing under which conditions we can hope to infer the shape of the individual infectiousness profile. We restrict to gamma-distributed timings, and consider cases where we assume that $\kappa$ is shared across individuals and where $\kappa$ comes from a beta distribution governed by hyperparameters $\alpha$ and $\beta$. Then, when $\kappa$ is identifiable, we introduce a hierarchical Bayesian inference strategy to infer the shape of the individual infectiousness profile given contact tracing data. 

For this section: let's imagine that there's a single $psi$ value shaed among everyone. We can use a Beta prior. Two things to consider: how do we estimate $\psi$ given a list of infection times (assume we know them exactly)? And how do we estimate $\psi$ given infection times and knowledge of $c_i(t)$? My hope is that this will emphasize that knowing the ``negatives'' -- i.e., the contacts that ``should'' have led to transmission but did not -- improves our inferential power substantially. Then, I want to think about how to achieve a given power in each scenario across R0, assuming some fixed generation interval distribution. That should be enough for now. We should probably include some amount of uncertainty in infection times for this base analysis, and we should post code for running the analysis given a line list of infection times and contacts. In the Discussion, we need to address the fact that there could be different $\psi_i$ values across individuals, themselves drawn from a distribution, so we'd want some kind of hierarhcical Bayesian model for inference here; also, we don't consider differences in the height (i.e. inherent biological differences in $\nu_i$ / $b_i$), which is another key thing to include in the inference. 