\clearpage 
\section{Introduction}

{\bf $A(\tau)$ is the fundamental epidemiological object.}
\begin{itemize}
	\item 	The trajectory of an epidemic depends on two fundamental quantities: the basic reproduction number ($R_0$), which is the expected number of secondary infections that an infectious person will produce in an otherwise susceptible population; and the intrinsic generation interval distribution ($g(\tau)$), which is the distribution of times at which those secondary infections are expected to occur relative to the index case's infection time ($\tau = 0$) \citep{wallinga_how_2007, champredon_intrinsic_2015}. 
	\item These quantities can be combined into a single object, the ``infectiousness profile'', defined as $A(\tau) = R_0 g(\tau)$, which describes a person's expected contribution to transmission and how it is distributed over time \citep{breda_formulation_2012}. This implies that $R_0 = \int_0^\infty A(\tau) \, d\tau$ and $g(\tau) = A(\tau) / R_0$. 
\end{itemize}

\noindent {\bf $A(\tau)$ is an expectation of $a_i(\tau)$.}
\begin{itemize}
	\item $R_0$ and $g(\tau)$ are both expectations taken over a large population. At the individual level, the individual reproduction number, $\nu_i$, is the expected number of infections that a particular person $i$ will produce, such that $R_0 = E_i[\nu_i]$ \citep{lloyd-smith_superspreading_2005}. Similarly, the ``individual infectiousness profile'' or ``infectiousness kernel'', $a_i(\tau)$, describes how a particular person's infectiousness is distributed over time, with $A(\tau) = E_i[a_i(\tau)]$ \citep{champredon_intrinsic_2015, svensson_note_2007}. 
	\item This implies the existence of an individual intrinsic generation interval distribution, $\xi_i(\tau)$, such that $a_i(\tau) = \nu_i \xi_i(\tau)$, with $\nu_i = \int_0^\infty a_i(\tau) d\tau$ and $\xi_i(\tau) = a_i(\tau) / \nu_i$. 
\end{itemize}

\noindent {\bf Since $A(\tau)$ is an expectation, it may look very different from $a_i(\tau)$.}
\begin{itemize}
	\item Since $A(\tau)$ is an expectation over individuals, it does not necessarily reflect any single person's infectiousness $a_i(\tau)$. 
	\item For example: the SEIR model is conceptually motivated by considering an individual who undergoes a latent period of length $x_i \sim \text{Exp}(\gamma)$ during which they are not infectious, followed by an infectious period of length $y_i \sim \text{Exp}(\alpha)$ during which they are infectious at level $\beta$. Mathematically, this ``stepwise'' individual infectiousness profile is expressed as
	%
	\begin{equation}
	a_i(\tau) = \begin{cases}
	    \beta & \tau \in [x_i, x_i + y_i] \\
	    0 & \text{otherwise}
	\end{cases}
	\label{eq:a_seir_stepwise}
	\end{equation}
	%
	This yields the population-level infectiousness kernel
	%
	\begin{equation}
	A_\text{{SEIR}}(\tau) = E_i[a_i(\tau)] = \lim_{n \rightarrow \infty} \frac{1}{n} \sum_{i=1}^n a_i(\tau) = \beta \frac{\gamma}{\gamma - \alpha}(e^{-\alpha \tau} - e^{-\gamma \tau})
	\label{eq:A_seir}
	\end{equation}
	%
	\ie, a smooth, bell-shaped curve that arises from a sum of discontinuous, stepwise $a_i(\tau)$ ({\bf Figure XX}). 
	\item Other choices of $a_i(\tau)$ can also yield the same $A_\text{{SEIR}}(\tau)$. For example, $a_i(\tau) \equiv A_\text{{SEIR}}(\tau)$ is a ``smooth'' individual infectiousness profile, where each person's infectiousness follows the population-level average. 
	\item Alternatively, $a_i(\tau) = \frac{\beta}{\alpha} \delta_{\tau_i^*}(\tau)$ is a ``spike'' individual infectiousness profile, where $\delta_{\tau_i^*}(\tau)$ is a delta function with mass at time $\tau_i^*$, and $\tau_i^*$ is distributed according to $g(\tau) = A_\text{{SEIR}}(\tau) / R_0$, so that a person's infectiousness is concentrated at a single moment. 
	% \item While all these choices yield the same mean-field population-level epidemic dynamics, since they all share the same $A(\tau)$, the potential impact on stochastic epidemic dynamics is poorly understood. 
	% \item Other choices of $a_i(\tau)$ can yield the same $A(\tau)$ in Eq.~\ref{eq:A_seir}, including
	% %
	% \begin{equation}
	% a_i(\tau) = \beta_{\text{SEIR}} \frac{\gamma_{\text{SEIR}}}{\gamma_{\text{SEIR}} - \alpha_{\text{SEIR}}} (e^{-\alpha_{\text{SEIR}} \tau} - e^{-\gamma_{\text{SEIR}} \tau})
	% \label{eq:a_seir_smooth}
	% \end{equation}
	% %
	% \ie, a ``smooth'' individual infectiousness profile where each person's infectiousness is identical to the population-level average, and
	% %
	% \begin{equation}
	% a_i(\tau) = \frac{\beta_{\text{SEIR}}}{\alpha_{\text{SEIR}}} \delta_{\tau_i^*}(\tau)
	% \label{eq:a_seir_spike}
	% \end{equation}
	% %
	% \ie, a ``spike'' infectiousness profile where $\delta_i(\tau)$ is a delta function with mass at time $t_i$, and $t_i$ is distributed according to the generation interval density $g(\tau) = A(\tau) / R_0$, such that a person's infectiousness is concentrated at a single moment ({\bf Figure \ref{fig:seir_sims} A--C}).
	% \item In general, any definition of $a_i(\tau)$ such that $E_i[a_i(\tau)] = A(\tau)$ will yield the same mean-field population-level epidemic dynamics; however, different choices of $a_i(\tau)$ may have important stochastic consequences ({\bf Figure \ref{fig:seir_sims} D--I}), which are the main topic of this article. 
\end{itemize}

\noindent {\bf We don't know what $a_i(\tau)$ looks like in general.} 
\begin{itemize}
	\item We have good methods for measuring $R_0$ and $g(\tau)$, but not for measuring $a_i(\tau)$. 
	\item Measuring $a_i(\tau)$ is challenging: it requires accurate contact tracing and precise inference of infection times. 
	\item We can maybe get at $a_i(\tau)$ from viral kinetics, but the link isn't obvious. Viral load $\neq$ viable virus shedding $\neq$ infectiousness. 
	\item The full variability in $A(\tau)$ is made up of both within-individual and between-individual variation: at one extreme, $a_i(\tau) \equiv A(\tau)$ (all variability in infection times comes from within-individual variation, and there is no between-individual variation), but this is unlikely to be true in general. Instead, individuals vary in the timing of infectiousness --- suggesting that $a_i(\tau)$ is in general narrower than $A(\tau)$. 
	\item Emphasizing this point, some evidence suggests that the period when SARS-CoV-2 and influenza can transmit explosively is short, lasting less than a day \citep{goyal_viral_2021}, suggesting that $a_i(\tau)$ may in some cases be much narrower than $A(\tau)$, with much of the variation in $A(\tau)$ arising from variation between individuals and on when their peak infectiousness occurs. 
\end{itemize} 

\noindent {\bf There are good reasons to believe that knowing about $a_i(\tau)$ is important.} 
\begin{itemize}
	\item This raises the question: if measuring $a_i(\tau)$ is so challenging, is there really any need to do it? Does it add anything beyond knowing $R_0$ and $g(\tau)$? There are good reasons to believe that it might. 
	\item At the individual level, communicating about the likely timing of a person's infectiousness is important: it is good to know if you will be very infectious for a short time, or mildly infectious for a long time, so that you can shape your behavior accordingly. 
	\item Additionally, a punctuated individual infectiousness kernel could impact epidemic dynamics. The stochastic effects of $a_i(\tau)$ early in an outbreak could cascade through the rest of the outbreak, creating meaningful differences in epidemic trajectory \citep{morris_computation_2024}. It could interplay with control measures, especially those designed to intervene on infectiousness at the individual level like test- or symptom-based screening and isolation. It could interact with time-varying contact patterns. 
	\item Indeed, the impact of the punctuatedness of the individual infectiousness kernel has been examined as a sensitivity analysis in various epidemiological studies \citep{larremore_test_2021} {\bf Others?}, but we lack a comprehensive understanding of its impact on epidemic dynamics. 
\end{itemize}

% \item However, we lack a comprehensive understanding of how the sharpness of the individual infectiousness kernel, independent of other features like the individual reproduction number, basic reproduction number, and generation interval distribution, impacts epidemic dynamics under both uncontrolled and controlled settings. 
% 	\item Furthermore, we lack an understanding of how to measure the shape of the individual infectiousness kernel reliably, along with the settings when the kernel's shape is statistically identifiable.

\noindent {\bf Our goals.}
\begin{itemize}
 	\item The main goal: holding $A(\tau)$ constant, how do different choices of $a_i(\tau)$ impact epidemic dynamics? And how would we know what $a(\tau)$ looks like? 
 	\item We'll focus on the punctuation of $a_i(\tau)$ as a key parameter, since it's perhaps the simplest and most plausible thing we can change. Effectively, this asks: how does the split of within-individual \vs between-individual variation in infection timing impact epidemic dynamics? 
 	\item We'll focus on pathogens with clear unimodal population-level and individual infectiousness kernels with short transmission times (\eg, respiratory viruses), though the theory can extend to pathogens with more complex transmission histories like tuberculosis and HIV. 
 	\item Interpersonal variation in the level of infectiousness has been studied extensively elsewhere \citep{lloyd-smith_superspreading_2005}, so we're going to focus on interpersonal variation in the timing of infectiousness (\ie, when it falls and how narrow the window of infectiousness is), holding all else constant, including critically the population-level parameters $R_0$ and $g(\tau)$. Thus, we'll be looking at stochastic effects around the mean-field solution.
 	\item We also address the question of how to infer $a_i(\tau)$ and examine its identifiability as a function of $R_0$. 
 \end{itemize} 

% ⏺ The literature search is done. Here's the key finding upfront: no paper makes your core results
%   redundant. But there are several important papers to cite and position against. Let me give you the
%    organized summary.

%   Papers you must cite and distinguish from

%   Most structurally similar:
%   - Yan (2008), J. Theor. Biol. — Decomposes the generation interval into latent + infectious periods
%    and studies how the split affects dynamics. Key difference: Yan changes A(τ) when varying the
%   split; you fix it. Also purely deterministic.
%   - Lloyd (2001a,b), Theor. Pop. Biol. & Proc. R. Soc. B — Shows narrower (Erlang) infectious periods
%    destabilize deterministic dynamics. Conceptually analogous (many E stages vs few I stages ≈
%   varying κ), but again changes A(τ), and is deterministic only.
%   - Wearing, Rohani & Keeling (2005), PLoS Medicine — Distribution shape of the infectious period
%   affects R0 estimates, peak size, intervention timing. Motivates your question but varies the
%   population-level profile.

%   Generation interval variance and growth rates:
%   - Wallinga & Lipsitch (2007), Proc. R. Soc. B — Foundational r-to-R0 mapping via GI shape. Their
%   "all GI equal to the mean" upper bound corresponds to your κ→0. Your project shows r and R0 are
%   κ-invariant by construction, but stochastic variability is not.
%   - Park, Champredon, Weitz & Dushoff (2019), Epidemics — Extends Wallinga-Lipsitch with Gamma GI
%   methods. Machinery you use, but no individual-level decomposition.

%   The W/time-shift framework:
%   - Morris, Maclean & Black (2024), J. Math. Biol. — Already cited. Provides the time-shift framework
%    but does not study how W depends on individual-level kernel decomposition or anything like κ.
%   - Jagers (1975), Nerman (1981), Jagers & Nerman (1984) — Foundational CMJ theory for W. Your Var(W)
%    as a function of κ extends their general results.

%   Superspreading / intervention papers:
%   - Lloyd-Smith et al. (2005), Nature — Foundational for overdispersion in offspring number. Does not
%    consider temporal variation. Your result that punctuated infectiousness + periodic contacts
%   generates overdispersion without R_i heterogeneity is a genuinely new mechanism.
%   - Goyal et al. (2021), eLife — Closest to your overdispersion mechanism (concentrated viral load +
%   heterogeneous contacts), but simulation-only, no analytical decomposition, no closed-form k(κ).
%   - Fraser, Riley, Anderson & Ferguson (2004), PNAS — Controllability depends on pre-symptomatic
%   transmission fraction θ. Your F_κ(δ_sym) generalises their θ by showing it depends on κ. They treat
%    it as a fixed pathogen property.
%   - Champredon & Dushoff (2015), Proc. R. Soc. B — Intrinsic vs realized generation intervals, but
%   focused on how dynamics filter the distribution, not on different individual kernels producing the
%   same population kernel.

%   What's genuinely novel in your project

%   1. The κ-parameterised family — a one-parameter family of individual infectiousness profiles that
%   all share the same population-level A(τ) via Gamma additivity. Nobody has formalised this
%   non-identifiability.
%   2. Var(W) as a function of κ — closed-form dependence of stochastic epidemic variability on the
%   individual-level decomposition, holding A(τ) and R0 fixed.
%   3. "Temporal superspreading" — punctuated infectiousness interacting with periodic contacts to
%   generate overdispersed offspring counts. The closed-form k(κ) is new.
%   4. Intervention effectiveness varying with κ — no paper examines how κ affects isolation/contact
%   tracing while holding A(τ) fixed.

%   Recommended follow-up searches

%   The agent couldn't access Google Scholar directly, so I'd suggest checking for 2025–2026 papers by
%   Morris's group, Park & Dushoff, Britton/Ball/Trapman, and Kucharski using terms like "clustered
%   transmission events," "correlated infection times branching process," "burstiness transmission,"
%   and "generation interval decomposition within between individual."